\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{multirow}
\usepackage{float}
\usepackage{caption}
\usepackage{tikz}
\captionsetup[table]{skip=10pt}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Case Study of Evaluating Porting Costs: Porting IxOS on ARM Boards}

\makeatletter
\newcommand{\linebreakand}{%
  \end{@IEEEauthorhalign}
  \hfill\mbox{}\par
  \mbox{}\hfill\begin{@IEEEauthorhalign}
}
\makeatother

\author{\IEEEauthorblockN{1\textsuperscript{st} Lucian-Ioan Popescu}
\IEEEauthorblockA{\textit{Computer Science and Engineering Department} \\
\textit{Politehnica University of Bucharest}\\
Bucharest, Romania \\
lucian\_ioan.popescu@stud.acs.upb.ro}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Lucian Mogosanu}
\IEEEauthorblockA{\textit{Research and Development} \\
\textit{Keysight Technologies}\\
Bucharest, Romania \\
lucian.mogosanu@keysight.com}
\linebreakand
\IEEEauthorblockN{3\textsuperscript{rd} Razvan Deaconescu}
\IEEEauthorblockA{\textit{Computer Science and Engineering Department} \\
\textit{Politehnica University of Bucharest}\\
Bucharest, Romania \\
razvan.deaconescu@cs.pub.ro}
}

\maketitle

\begin{abstract}
This document is a model and instructions for \LaTeX.
This and the IEEEtran.cls file define the components of your paper [title, text, heads, etc.]. *CRITICAL: Do Not Use Symbols, Special Characters, Footnotes,
or Math in Paper Title or Abstract.
\end{abstract}

\begin{IEEEkeywords}
component, formatting, style, styling, insert
\end{IEEEkeywords}

\input{introduction}
\input{background}
\input{porting-ixos}
\input{evaluation}
\input{discussions}
\input{conclusions-and-further-work}

\begin{thebibliography}{00}
\bibitem{b1} Tanaka, T.; Hakuta, M.; Iwata, N.; Ohminami, M. (1995). [IEEE Comput. Soc. Press 12th TRON Project International Symposium - Tokyo, Japan (28 Nov.-2 Dec. 1995)] Proceedings of the 12th TRON Project International Symposium - Approaches to making software porting more productive. , (), 73–85.
\bibitem{b2} Mitsuari Hakuta; Masato Ohminami (1997). A study of software portability evaluation. , 38(2), 145–154.
\bibitem{b3} L. Fernando Capretz, "Bringing the Human Factor to Software Engineering" in IEEE Software, vol. 31, no. 02, pp. 104-104, 2014.
\bibitem{b4} A. Kanai, T. Furuyama and M. Takahashi, "A cost model for software conversion based on program characteristics and a converter effect," [1992] Proceedings. The Sixteenth Annual International Computer Software and Applications Conference, 1992, pp. 63-68
\bibitem{b5} Mooney, J.D., 2004. Developing portable software. In Information Technology (pp. 55-84). Springer, Boston, MA.
\bibitem{b6} Ejiogu, L.O., 1985. A simple measure of software complexity. ACM SIGPLAN Notices, 20(3), pp.16-31.
\end{thebibliography}

\end{document}
