\chapter{Introduction}\pagestyle{fancy}

Porting a software system and evaluating the porting costs is a hard problem in
systems
programming.~\cite{tanaka,hakuta,mooney2004developing,kanai,porquet2015,
bodenstab1984unix,osdevcrossport,jolitz1990porting,frakes1995sixteen,
tanenbaum1978guidelines,johnson1978unix}
The reasons for porting software systems are various: the developers want to
enhance the performance, the hardware environment starts to get deprecated or
the system wants to take advantage of features unavailable in the current
environment.

Software development costs and invested resources for understanding, maintaining
and developing new features for a system are
big~\cite{xia2017measuring,boehm2000software,morgan1994controlling}. We~\lm{you use ``we'', but then you suddenly switch to ``people''} would
like to preserve these investments when the need for a new environment arises.
To achieve this goal people designed programming languages and compilers that
increased portability, operating systems that could run on multiple hardware
platforms~\cite{johnson1978unix} and various standards that allowed developers to talk to
the computer using well defined interfaces~\cite{walli1995posix}. However the porting
process is a non-trivial endeavor up to this day, it is error prone and time
consuming.

As with other software development processes, porting has its own costs
associated. It is important to evaluate them and understand their implications
in the project so that the developer can optimize the process of porting in the
future and call attention to the weaknesses and strengths of the process. In
this work we describe the experience of porting the IxOS testing infrastructure,
used by Ixia for high performance network testing, on ARM off-the-shelf boards.
We do this for two reasons. Firstly, we are interested in exploring Ixia testing
infrastructure on newer environments with the hope that we will improve the
performance and reduce the costs of delivering the system. Secondly, we are
interested in evaluating porting costs related to the project for the reasons
listed in the beginning of this paragraph.~\lm{This last sentence sounds a bit awkward; you can rephrase the paragraph by saying directly that you want to understand the costs involve in porting; and that as an experimental vehicle for this, you choose Ixia testing software, which is known to be industry-grade and versatile, and thus is sufficiently complex as a practical case study.}

On one hand, we make the following contributions in this work: we review the
porting models described in~\cite{tanaka,hakuta,kanai} and propose a more
general model that can be applied to modern software porting, we review the
porting costs factors described in~\cite{hakuta} and analyze their relevance in
our porting work, and we provide guidelines and discussions on the topic of
improving software porting based on the experience of porting IxOS
infrastructure. On the other hand, we ported~\lm{you suddenly switch to past tense here; since you're in the intro, I suggest leaving the present tense.} the testing infrastructure on ARM
arhitecture. During this process we isolated the relevant parts of our porting
in a separate repository. As a result, the isolated components can now be used
independently from IxOS by internal teams interested in further developing of
testing tools and systems on ARM boards.

This paper~\lm{thesis?} highlights the porting issues associated with porting the IxOS
network testing infrastructure, both the hardware and software components, to
ARM off-the-shelf boards, Raspberry Pi 4 in our case~\lm{It would be clearer if you started this paragraph off with something like: ``This thesis is structured as follows. In Chapter 1 etc.''}. In
Chapter~\ref{sec:background} we present various terms associated with porting:
what porting is, porting models, porting costs and factors. In
Chapter~\ref{sec:portingIxos} we divide our work~\lm{try to say in two words what work exactly you're referring to, i.e. you're actually presenting the porting architecture and steps.} in multiple steps. For each
step we present a description, the targets and milestones, and the final
results. After the porting process is discussed, we highlight in
Chapter~\ref{sec:eval} the porting costs associated with our work, including an
analysis of the factors that influenced the presented costs. Finally, in
Chapter~\ref{sec:discussion} we conduct a discussion about the porting process
and the porting costs where we investigate porting difficulties, observations
about the porting tasks and ways to improve the porting process.~\lm{It's also customary to say something about the concluding chapter, just so that the reader know it's there.}
